### 2014-03-11 paper.Rnw

% !Rnw weave = knitr
\documentclass{article}

\usepackage{cite}
\begin{document}

<<setup, include=FALSE, cache=FALSE, echo=FALSE>>=
Sys.setenv(TEXINPUTS=getwd(), 
           BIBINPUTS=getwd(), 
           BSTINPUTS=getwd())
@

\title{Analysis of Massachusetts Standardized Testing Data through Structural Equation Modelling}
\date{March 24, 2014}
\author{Landon Hurley\\ Psychology, SUNY Geneseo \and Herb Susmann\\ Mathematics, SUNY Geneseo}

\maketitle

\section*{Abstract}

\section*{Introduction}

\section*{Data}

The Massachusetts Compreshensive Assessment System (MCAS) is a standardized test administered to Massachusetts public school students in grades 3-10 since 1998. We examine a random sample of 10,515 10th grade student's results from the 2009 MCAS test. The data also contains basic demographic information for each student.

The test consists of english, mathematics and science components. Every student takes a English and Mathematics tests; the science component can be fulfilled by taking either a Biology, Chemistry, Introductory Physics, or Technology/Engineering test. \cite{mcas_summary}

Each test consists of multiple choice and short answer type questions.

\section*{Methods}
Description of SEM

\section*{Results}

\section*{Discussion}
Could also include ideas for other questions to ask in the questionairre

\section*{Works Cited}

\bibliography{references}{}
\bibliographystyle{plain}

\end{document}