### 2014-04-09 presentation.Rnw
Presentation beamer file

\documentclass{beamer}

\title{Analysis of Massachusetts' Standardized Testing Multi-Group Structural Equation Modelling}
\author[shortname]{Landon Hurley \inst{1} \and Herb Susmann \inst{2}}
\institute[shortinst]{\inst{1} Psychology, SUNY Geneseo \and %
\inst{2} Mathematics, SUNY Geneseo}
\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}
  \tableofcontents
\end{frame}

\begin{frame}{Introduction}

\end{frame}

\section{Data}

  \begin{frame}{Data}
    \begin{itemize}

    \end{itemize}
  \end{frame}

\section{Measurement Invariance}

  \begin{frame}{Measurement Invariance}

  \end{frame}

  \subsection{Results}

    \begin{frame}{Results}

    \end{frame}

  \subsection{Discussion}

    \begin{frame}{Discussion}

    \end{frame}

\section{Regression Tree}
  \begin{frame}{Regression Tree}

  \end{frame}

  \subsection{Results}

    \begin{frame}{Results}

    \end{frame}

  \subsection{Discussion}

    \begin{frame}{Discussion}

    \end{frame}

\end{document}